Global patterns of forest autotrophic carbon fluxes

### 2020-03-01 manuscript/manuscript draft.Rmd
worked on results

This analysis focused on nine FACF included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]. ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. Associated data, such as stand age, measurement methodologies, and disturbance history, are also included. The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010]. Additional targeted literature searches were conducted to identify any further available data on the FACF analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016. We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.

@@

All major carbon fluxes decreased linearly with latitude (Fig. 2).

Latitude was a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). For all fluxes, their relationship with latitude was best predicted by the linear model.

*Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (fig. 2). That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.

bbl: here or in the discussion, note that this is a fairly stringent test: it's easy for sub-fluxes not to sum up! (Extensive citations from EC literature for example.). HML: yes and no. confidence intervals are pretty large.

@@

*We found no evidence that allocation between fluxes varied substantially with latitude or climate. There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.* **THIS HAS CHANGED: [issue #51](https://github.com/forc-db/Global_Productivity/issues/51#issuecomment-581892186)

>>>>

This analysis focused on nine FACF included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]. ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. Associated data, such as stand age, measurement methodologies, and disturbance history, are also included. The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010]. Additional targeted literature searches were conducted to identify any further available data on the FACF analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016. We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855. This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.

@@

All major carbon fluxes decreased linearly with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$=0.18). For all FACF, the relationship with latitude was best predicted by the linear model (**but see [this issue]((https://github.com/forc-db/Global_Productivity/issues/64))**).

@@

In general, smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (Fig. 2). That is, modelled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{woody-stem}$, fell completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.

We found little evidence that allocation between fluxes varied substantially with latitude or climate (Fig. [**S3**](https://github.com/forc-db/Global_Productivity/issues/65)). Of the 7 FACF ratios regressed against latitude and three climate variables (MAT, MAP, temperature seasonality), there were only four signficant relationships, all with $R^2 \le 0.2$ (Fig. S3). Specifically, the proportion of NPP allocated to $ANPP_{foliage}$ increased weakly with MAT ($R^2=0.20$), and the proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).  There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.

(*KAT has revised results up to here.*)
